Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/design.tex
Fixed citations
Removed figure 2(DNS example) Re-wrote dependent fields

The \emph{many} combinator takes a parser and applies it repeatedly
until it fails, returning an array of the inner parsers results. The
\emph{sepBy} combinator
additionally takes a constant parser, which it applies in between parsing
two values, but not before parsing the first value or after parsing the
last.
For example, \cc{many uint8} represents an array of 8-bit unsigned
integers, and \cc{sepBy uint8=',' (many uint8 | '0'..'9')} recognizes
comma-separated lists of decimal numbers.

@@

\begin{figure}[tb]
\smaller[0.5]
\begin{minipage}{0.45\columnwidth}
\begin{verbatim}
dns_header = {
  id uint16
  qr uint1
  opcode uint4
  aa uint1 
  tc uint1
  rd uint1
  ra uint1
  uint3 = 0
  rcode uint4
}
\end{verbatim} 
\end{minipage}
~
\begin{minipage}{0.45\columnwidth}
\begin{verbatim}
struct dns_header {
  uint16_t id;
  uint8_t qr;
  uint8_t opcode;
  uint8_t aa;
  uint8_t tc;
  uint8_t rd;
  uint8_t ra;
  uint8_t rcode;
};
\end{verbatim} 
\end{minipage}

\caption{Nail grammar (left) and data model (right) for a part of the
grammar for DNS packets.  The \cc{uint3 = 0} grammar represents
3 bits of padding (filled with zeroes).}
\label{fig:dns-struct}
\end{figure}

@@

A sample grammar for a DNS header and the corresponding C data type the Nail parser produces, are
shown in Figure~\ref{fig:dns-struct}.

@@

As discussed above, choosing to use these combinators on constant parsers
removes the bijection between byte-strings and our data model, as there are
multiple byte-strings that correspond to the same internal data structure and
the generator has to choose one of these representations.
\paragraph{Wrap combinator.} 
When implementing real protocols with Nail, we often found that
structures that consist of many constant parsers and only one named field. This pattern is
common in binary protocols which use fixed headers to denote the type of data
structure to be parsed.  In order to keep the internal representation cleaner,
we introduced the wrap combinator, which takes a sequence of parsers containing
exactly one non-constant parser. The parser and generator act as though the wrap
combinator was a sequence of parsers, but the data model does not wrap the
single value in another structure, making the application-visible representation
(and thus application code) more concise.

@@

Dependent fields are defined within a structure like normal fields, but their name starts with an \@-sign. 
A dependent field is in scope and can be referred to by the definition of all subsequent fields in
the same structure. Dependency fields can be passed to rule invocations as parameters.

During parsing, dependent fields are parsed just like other fields, but instead of storing their
value in a structure that will be returned to the programmer, the parser stores the value in a
temporary location. Thereafter, the value can be referenced in the grammar.

When generating output, Nail visits a dependent field twice. When generating the fields of a
structure, it will reserve space for the dependent field in the output, but not write a value. When
the dependent field goes out of scope, the generator will backtrack to this space and write the
dependent fields value.

>>>>

The \emph{many} combinator takes a parser and applies it repeatedly
until it fails, returning an array of the inner parsers results. The
\emph{sepBy} combinator
additionally takes a constant parser, which it applies in between parsing
two values, but not before parsing the first value or after parsing the
last.
%For examle, \cc{many uint8} represents an array of 8-bit unsigned
%integers, and \cc{sepBy uint8=',' (many uint8 | '0'..'9')} recognizes
%comma-separated lists of decimal numbers.

@@

% \begin{figure}[tb]
% \smaller[0.5]
% \begin{minipage}{0.45\columnwidth}
% \begin{verbatim}
% dns_header = {
%   id uint16
%   qr uint1
%   opcode uint4
%   aa uint1 
%   tc uint1
%   rd uint1
%   ra uint1
%   uint3 = 0
%   rcode uint4
% }
% \end{verbatim} 
% \end{minipage}
% ~
% \begin{minipage}{0.45\columnwidth}
% \begin{verbatim}
% struct dns_header {
%   uint16_t id;
%   uint8_t qr;
%   uint8_t opcode;
%   uint8_t aa;
%   uint8_t tc;
%   uint8_t rd;
%   uint8_t ra;
%   uint8_t rcode;
% };
% \end{verbatim} 
% \end{minipage}

% \caption{Nail grammar (left) and data model (right) for a part of the
% grammar for DNS packets.  The \cc{uint3 = 0} grammar represents
% 3 bits of padding (filled with zeroes).}
% \label{fig:dns-struct}
% \end{figure}

@@

% A sample grammar for a DNS header and the corresponding C data type the Nail parser produces, are
% shown in Figure~\ref{fig:dns-struct}.

@@

As discussed above, choosing to use these combinators on constant parsers
weakens the bijection between the format and the data type, as there are
multiple byte-strings that correspond to the same internal representation and
the generator chooses one of these.
\paragraph{Wrap combinator.} 
When implementing real protocols with Nail, we often found that
structures that consist of many constant parsers and only one named field. This pattern is
common in binary protocols which use fixed headers to denote the type of data
structure to be parsed.  In order to keep the internal representation cleaner,
we introduced the wrap combinator, which takes a sequence of parsers containing
exactly one non-constant parser. The parser and generator act as though the wrap
combinator was a sequence of parsers, but the data model does not wrap the
single value in another structure, making the application-visible representation
(and thus application code) more concise.

@@

Dependent fields are defined within a structure like normal fields, but their name starts with an $@$-sign. 
A dependent field is in scope and can be referred to by the definition of all subsequent fields in
the same structure. Dependency fields can be passed to rule invocations as parameters.

Dependent fields are parsed just like other fields in the external format, but they are not stored
in the internal data type. Instead, the parser stores the value in a
temporary location the field can be referenced by subsequent parsers.

When generating output, Nail visits a dependent field twice. First, while generating the other fields of a
structure, the generator reserves space for the dependent field in the output. Once
the dependent field goes out of scope, the generator  write the
dependent field's value to this space.
